
\section{Digital Economy}

Digital economy can be thought of as the economy that is based on the internet and digital technologies.

\begin{definition}
    {Types of digital economy}
    \begin{itemize}
        \item \textbf{E-commerce}: Digital enabling of commercial transactions
              \\      \textit{Examples: Online shopping, mobile banking etc.}
        \item \textbf{E-business}: The use of the internet to conduct business activities, such as marketing, sales, and customer service.
              \\      \textit{Examples: Online marketing, online sales, online customer service etc.}
    \end{itemize}
\end{definition}

\subsection{Key characteristics}

\subsubsection{Data integration to the product process}

The traditional production process is as followed:

Design\quad\rightarrowfill\quad
Raw Material\quad\rightarrowfill\quad
Manufacturing\quad\rightarrowfill\quad
Finished Product\quad\rightarrowfill\quad
Consumer

We can \textbf{monetize data} that we collect from the \textit{consumers} by analyzing it and apply in the processes during production.

\subsubsection{Digital products}

\begin{definition}
    {Digital products}
    Intangible goods that exist in digital forms. \textit{Examples: Software, music, databases, platforms \& services etc.}
\end{definition}

A digital good has the following distinctive characteristics:

\begin{itemize}
    \item \textbf{Zero marginal cost}: There is no cost to produce additional units of a digital good.
    \item \textbf{Quality uncertainty}: Consumers cannot determine the value of digital goods before using it.
\end{itemize}

Some digital goods also increase in value the more people use it, known as \textbf{network goods}.

\begin{definition}
    {Types of network goods}
    \begin{itemize}
        \item \textbf{Homogeneous}: Users perform the same action. (e.g. Whatsapp)
        \item \textbf{Two-side}: Distinct user groups (e.g. Carousell)
    \end{itemize}
    The types of goods are usually \textbf{platforms}, and it's the role of the \textbf{platform owner} to design and manage the playfield.
\end{definition}